### 2022-02-03 _posts/welcome/welcome.Rmd
2.3.22
1. Recreating 679 SNA Assignment Blog
2. Create GitRepro & Distill Blog
Linberg (2022, Jan. 29). slinberg.net: Distill and Github Pages Setup. Retrieved from https://slinberg.net/meta/distill-and-github-pages-setup/

3. Setup- Publishing to RPubs

---
title: "Welcome to Test: Distill Blog"
description: |
  Welcome to our new blog, Test: Distill Blog. We hope you enjoy 
  reading what we have to say!
author:
  - name: AJW
    url: https://example.com/norajones
    affiliation: Spacely Sprockets
    affiliation_url: https://example.com/spacelysprokets
date: 02-03-2022
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Learn more about creating blogs with Distill at:
# https://rstudio.github.io/distill/blog.html

```