### 2022-02-03 SetUp_RPubs_PublishBlog.Rmd

---
title: "SetUp_RPubs_PublishBlog"
description: |
  Publishing to RPubs: A new article created using the Distill format.
author:
  - name: 
    url: 
    affiliation: 
    affiliation_url:
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

```

*Publishing to RPubs*

Open RStudio and create a new Rmd file
                            (File > New File > R Markdown).
When you’re ready to publish to RPubs, click Knit.
From your knitted document, click Publish in the top, right-hand corner.
Select Rpubs.

Click Publish. You’ll then be able to create/edit a title, short description and customized slug.

********
Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill for R Markdown at <https://rstudio.github.io/distill>.