### 2022-02-08 docs/1.SNA-  Templates/SNA-Template_Weekly-Assignments.Rmd

---
title: "Social Network Analysis: Assignment: Week#"
description: " lorum ipsum "
  A new article created using the Distill format.
author:
  - name: "Audra Jamai White" 
    url: 
    affiliation: " DACSS 679: Social Network Analysis"
    affiliation_url: 
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, Libraries, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Document Creation
library(knitr)
library(rmarkdown)
library(yaml)
library(distill)
library(blogdown)

# Basic Analysis
library(ggplot2)
library(tidyverse)
library(clipr)
library(dbplyr)
library(data.table)
library(stringr)
library(stringi)
library(tibble)
library(tidyr)

# SNA - Social Network Analysis

library(tergm)
library(ergm)

library(network)
library(networkDynamic)
library(networkDynamicData)

library(statnet)
library(igraph)
library(igraphdata)
library(tsna)
library(sna)

```

*Week \# Assignment*

```{r,Import Data Set,include=FALSE}

```

```{r, Describe Data Set}

```

```{r, Analyze Data Set}

```

Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill for R Markdown at <https://rstudio.github.io/distill>.