### 2022-02-08 docs/2. SNA-Tutorials/week2-SNA_Tutorial.Rmd

---
title: "SNA Tutorials: Week 2: Network Structure"
output: html_notebook
---

**[igraph]**
```
```

**[statnet]**
```
```

## Key Terms & Definitions
### Vetex Attribute & Edge Attribute 
A *vertex attribute* is a vector (or variable) of additional information about the nodes in a network dataset:
name, location, demographic information about the actors or organizations  comprise the vertices of the network.

In case you have forgotten, we can 
[statnet]
``` print(DataSet.stat) 
```

*edge attribute* is a vector (or variable) of additional information about the edges or relationships in a network dataset. The most common edge attribute is weight, in a weighted dataset.
### Network Structure
- Dyads and Dyad Census
  *dyad census* gives the number of dyads where the relationship is:
    - Reciprocal (mutual), or mut
    - Asymmetric (non-mutual), or asym, and
    - Absent, or null

- Triads and Triad Census
  *triad census* provides a fundamental descriptive insight into the types of triads found in a dataset

- Network Transitivity / Global Clustering
  *Transitivity* is a network-level statistic that captures the proption of connected triads in the network that are complete (all three possible links present)
  *Transitivity* varies between 0 [no connected triads] and 1 [all connected triads are transitive]

- Local Transivity or Clustering
  Local transitivity (*local clustering coefficient*) is the density of an ego network
  *average clustering coefficient* average of local transivity for all vertices in a graph

- Path Length & Geodesic

- Component Structure and Membership

# Network Structure [Week 2]
aspects of network structure introduced in the Week 2 Lecture:
- Dyads and Dyad Census
  *dyad census* gives the number of dyads where the relationship is:
    - Reciprocal (mutual), or mut
    - Asymmetric (non-mutual), or asym, and
    - Absent, or null

- Triads and Triad Census
  *triad census* provides a fundamental descriptive insight into the types of triads found in a dataset

- Network Transitivity / Global Clustering
  *Transitivity* is a network-level statistic that captures the proption of connected triads in the network that are complete (all three possible links present)
  *Transitivity* varies between 0 [no connected triads] and 1 [all connected triads are transitive]

- Local Transivity or Clustering
  Local transitivity (*local clustering coefficient*) is the density of an ego network
  *average clustering coefficient* average of local transivity for all vertices in a graph

- Path Length & Geodesic

- Component Structure and Membership

# Describe Network Data 
## 1. *List and inspect:* 
    List the objects to make sure the datafiles are working properly:
```is()
```

## 2. *Network Size:*
    What is the size of the network?

**[igraph]**
    Count of vertices
```vcount(DataSet.ig)
```
    Count of edges
``` ecount(DataSet.ig)
```
**[statnet]**
    Network Attributes: count vertices and edges 
```print(DataSet.stat)
```

## 3. *Network Features:*

  **[igraph]**    
  Is this a Bipartite or single mode network?
``` is_bipartite(DataSet.ig)
```

  Are edges directed or undirected?
```is_directed(DataSet.ig)
```

  Are edges weighted or unweighted?
```is_weighted(DataSet.ig)
```

  **[statnet]**
    Find network features
```print(DataSet.stat)
```

## 4. *Network Attributes*
  R commands used with vector objects also work with network attribute data
```sort()
```
```summary()
```

**[igraph]**
```summary(
      E(
      DataSet.ig
      )
      $weight
    )
 ```

**[statnet]**
``` summary(
      DataSet.stat 
      %e% 
      "weight"
    )
```

  *Accessing Attribute Data*
- When we work with an R dataframe object, you can access a specific variable in the dataframe by using the >data$varname< or >$ command<

  **[igraph]**
  access vertex attribute 
```V(
      DataSet.ig
      )
      $name
```
  access edge attribute
```E(
      DataSet.ig
      )
      $weight
```

  **[statnet]**
  access vertex attribute
```head(
      DataSet.stat 
      %v% 
      "vertex.names"
      )
```
  access edge attribute
``` head(
        DataSet.stat 
        %e% 
        "weight"
        )
```

  Generate a list the names of vertex and edge attributes of the network data.
  **[igraph]**
  Names- Vertex Attributes
```igraph::vertex_attr_names(climate.ig)
```
  Names- Edge Attributes
```igraph::edge_attr_names(climate.ig)
````

  **[statnet]**
  Names- Vertex Attributes
```network::list.vertex.attributes(climate.stat)
```
  Names- Edge Attributes
```network::list.edge.attributes(climate.stat)
```

# Dyad Census 
  Classifies all dyads in the network as:
      - Reciprocal (mutual), or mut
      - Asymmetric (non-mutual), or asym, and
      - Absent, or null

  **[igraph]**
```igraph::dyad.census(gotbook.ig)
```  
  **[statnet]**
```sna::dyad.census(gotbook.stat
```

# Triad Census
  Classifies all triads in the network

**[igraph]**
``` igraph::triad_census(gotbook.ig)
```
**[statnet]**
  Undirected Network
``` sna::triad.census(gotbook.stat, mode="graph")```
  Directed Detwork
``` sna::triad.census(gotbook.stat, mode=)```

# Transitivity or Global Clustering

  network transitivity:
**[igraph]**
``` transitivity(gotbook.ig)
```
  weighted network transitivity:
```igraph::transivity()
```

**[statnet]**
  weighted network transitivity:
``` sna::gtrans()
```
  For Directed Networks - network structure rooted in hierarchy
```gtrans()
```
# Local Transivity or Clustering

  **[igraph]**

  1. Retrieve a list of the three vertices we are interested

``` V(gotbook.ig)
      [c(
      "vertex_attr_names",
      "vertex_attr_names",
      "vertex_attr_names"
      )]
```

  2. Check ego network transitivity

```transitivity(
      gotbook.ig,
      type="local",
      vids=V(gotbook.ig)
      [c(
      "vertex_attr_names",
      "vertex_attr_names",
      "vertex_attr_names"
      )]
    ) 
```

  Global clustering cofficient
```transitivity(gotbook.ig,
      type="global"
    )
```

  Average local clustering coefficient
```transitivity(gotbook.ig,
      type="average"
    )
```

# Advanced [Week 2]
## ADVANCED: Path Length and Geodesic
## ADVANCED Component Structure and Membership