### 2022-02-08 docs/3. SNA - Assignments/week2-Assignment-Basic_Network_Structure.Rmd

---
title: "Social Network Analysis: Week 2: Basic Network Structure"
description: "use `igraph` and `statnet` tools to describe aspects of network structure introduced in the Week 2 Lecture: Dyads and Dyad Census, Triads and Triad Census, Network Transitivity and Clustering, Path Length & Geodesic"

author:
  - name: "Audra Jamai White" 
    url: 
    affiliation: "Umass Amherst - DACSS 679: Social Network Analysis"
    affiliation_url: 
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, Libraries, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Document Creation
library(knitr)
library(markdown)
library(yaml)
library(distill)
library(blogdown)

# Basic Analysis
library(ggplot2)
library(tidyverse)
library(clipr)
library(dbplyr)
library(data.table)
library(stringr)
library(stringi)
library(tibble)
library(tidyr)

# SNA - Social Network Analysis

library(tergm)
library(ergm)
library(network)
library(networkDynamic)
library(networkDynamicData)

library(statnet)
library(igraph)
library(igraphdata)
library(tsna)
library(sna)

```

## Social Network Analysis

# **Week 2 Assignment: Network Structure**

1.  [**Identify an existing data set.**]{.ul}

| This can be one provided in the course directory, in an R package or library, located online, or some other source.

2.  [**Briefly describe the network dataset.**]{.ul}

| Identify and describe content of nodes and links, and identify format of data set (i.e., matrix or edgelist, directed or not, weighted or not), and whether attribute data are present. 
  Be sure to provide information about network size (e.g., information obtained from network description using week 1 network basics tutorial commands.)

3.  [**Explore the dataset using commands from week 2 tutorial.**]{.ul}

| Comment on the highlighted aspects of network structure such as geodesic and path distances, triads or transitivity, connectedness and.or component structure, etc. 
  Be sure to both provide the relevant statistics calculated in R, as well as your own interpretation of these statistics.
|

# 1. Identify an existing data set.

-   **Data set:** Florentine Families

-   **Description:** This is an adjacency matrix for a small, directed network of the marriages held by the Florentine family and others in Rome.

-   **Package:** [network](https://cran.r-project.org/web/packages/network/network.pdf)

-   **Access:**

    -   [Import Script]{.ul}:

        `>source("./Import Scripts/Florentine Families.R")`

    -   [Orignial Format from Package]{.ul}:

        `>install.packages("network")`

        `>data("flo", package = "network")`

```{r Import Data Set, echo=TRUE}

  # install.packages("network")

    data("flo", package = "network")

```

# 2. Briefly describe the network dataset.

-   *Be sure to provide information about network size (e.g., information obtained from network description using week 1 network basics tutorial commands.)*

1.  *Identify and describe content of nodes and links,*

-   each of the 16 Florentine Family Names represent a Node in the network.

```{r Colnames}
 colnames(flo)
```

2.  *Identify format of data set (i.e., matrix or edgelist, directed or not, weighted or not), and*

    -   Flo Format: Size: 16 x 16 square dimmentions indicate a adjacency matrix format

    Network attributes: vertices = 16 directed = TRUE hyper = FALSE loops = FALSE multiple = FALSE bipartite = FALSE total edges= 40 missing edges= 0 non-missing edges= 40

Vertex attribute names: vertex.names

No edge attributes

```{r Dimmentions, echo=TRUE}

  dim(flo)

```

```{r Network Attributes, echo=TRUE}

  network(flo)

```

3.  *Identify whether attribute data are present.*

```{r Summary,echo=TRUE}

  summary(flo,
          print.adj = TRUE
          )
```

## Create & Describe: Directed Adjacency Matrix

*Florentine Families Directed Adjacency Matrix Network Objects:*
    **iGraph** flo.ig 
    **Statnet** flo.stat

iGraph and Statnet both use {r graph4} for Directed Adjacency Matrix Objects

```{r graph4, statnet: Directed: Adjacency Network, echo=TRUE}

    flo.stat<-
        network(
        flo,
        directed=F,
        matrix.type="adjacency"
        )

#     provides a description of several critical network features
    print(flo.stat)

```

```{r graph4, igraph: Directed: Adjacency Network, echo=TRUE}

    flo.ig<-
      graph.adjacency(
      flo,
      mode="undirected"
      ) 
    print(flo.ig)
```

```{r igraph: Network Description, echo=TRUE}

#     Count Vertices
    vcount(flo.ig)

#     Count Edges
    ecount(flo.ig)

# Is this a Bipartite or single mode network?
    is_bipartite(flo.ig)

#  Are edges directed or undirected?
    is_directed(flo.ig)

#Are edges weighted or unweighted?
    is_weighted(flo.ig)

```
 **Vertex and Edge Attributes**

```{r statnet: Attribute Data: Directed: Adjacency Network, echo=TRUE}

#     access vertex attributes
  head(flo.stat %v% "vertex.names")

#      list the names of vertex attributes
  network::list.vertex.attributes(flo.stat)

#     access edge attribute
  head(flo.stat%e% "weight")

#  Generate a list the names of edge attributes
  network::list.edge.attributes(flo.stat)

#   summarize numeric network attribute
  summary(flo.stat  %v% "name")

```

```{r igraph: Attribute Data: Directed: Adjacency Network, echo=TRUE}

#     access vertex attribute 
  V(flo.ig)$name
#     Generate a list the names of vertex attributes
  igraph::vertex_attr_names(flo.ig)

#     access edge attribute
  E(flo.ig)$weight

#   Generate a list the names of edge attributes  
  igraph::edge_attr_names(flo.ig)

#   summarize numeric network attribute
  summary(E(flo.ig)$weight)

```

# 3. Explore the dataset using commands from week 2 tutorial.

5.  *geodesic and path distances,*

## Path Length and Geodesic

```{r iGraph:Nodes: Path: Length, echo=TRUE}

#     Calculate distances between two nodes
    distances(flo.ig,"Bischeri","Ridolfi")

#     Calculate distance between two nodes using unweighted edges
    distances(flo.ig,"Bischeri", "Castellani",weights=NA)

```

We can also find all of the shortest paths between two famalies in this network.
These list the starting node and the ending node with all nodes inbetween for each path. 
The option weights=NA means that any available edge weights are ignored.

```{r iGraph:Nodes: Path: Shortest, echo=TRUE}

#     isolating the distances between specific node
    all_shortest_paths(flo.ig,"Strozzi","Tornabuoni", weights=NA)$res
## Note: manually tell igraph to ignore edge weights.
```

The concept of shortest path to describe the overall network structure can be more useful

```{r iGraph:Network: Path: Shortest, echo=TRUE}

#     find average shortest path for network
    average.path.length(flo.ig,directed=F)

```

6.  *triads or transitivity,*

## Dyad Census

Classifies all dyads in the network as: - Reciprocal (mutual), or mut - Asymmetric (non-mutual), or asym, and - Absent, or null

```{r statnet: Dyad Census , echo=TRUE}

    sna::dyad.census(flo.stat)

```

```{r iGraph: Dyad Census, echo=TRUE}

  igraph::dyad.census(flo.ig)

```

## Triad Census

Classifies all triads in the network. The triad census provides a fundamental descriptive insight into the types of triads found in a particular dataset.

```{r statnet: Triad Census, echo=TRUE}

#     Classify all Triad in the network: Directed Network
  sna::triad.census(flo.stat, mode= )

#     Classify all Triad in the network: Undirected Network
   sna::triad.census(flo.stat, mode="graph")

#     total number of all four triad types returned by triad census
   sum(sna::triad.census(flo.stat, mode="graph"))

```

```{r iGraph: Triad Census, echo=TRUE}
#Classify all Triad in the network
     igraph::triad_census(flo.ig)

```

## Transitivity or Global Clustering

```{r iGraph: weighted/ network transitivity, echo=TRUE}

#     network transitivity:
    transitivity(flo.ig)

#     weighted network transitivity:
#   igraph::transivity(flo.ig)

```

```{r statnet: weighted/ directed network transitivity, echo=TRUE}

#     Directed Networks transitivity:
  gtrans(flo.stat)

#   methods for weighted networks, rank and correlation, along with the relevant references.
#  sna::gtrans()

```

## Local Transivity or Clustering
`Local transitivity` (local clustering coefficient), is a technical description of the density of an ego network.

```{r iGraph:Local transitivity, echo=TRUE}

#     Retrive a list of the vertices we are interested
    V(flo.ig)[c("Bischeri" ,
                "Castellani", 
                "Medici",
                "Ridolfi")]

#     check ego network transitivity
transitivity(flo.ig,
             type="local", 
             vids=V(flo.ig)
             [c("Bischeri",
                "Castellani", 
                "Medici",
                "Ridolfi")]) 
#     get global clustering cofficient
  transitivity(flo.ig, type="global")

#     get average local clustering coefficient
  transitivity(flo.ig, type="average")

```

7.  *connectedness and.or component structure, etc.*

> Distill is a publication format for scientific and technical writing, native to the web.
>
> Learn more about using Distill for R Markdown at <https://rstudio.github.io/distill>.