### 2022-02-19 docs/3. SNA - Assignments/Week3-Distill-Blog_Degree-and-Centrality.Rmd
11.??.22
I took a break

---
title: "Social Network Analysis: Week 3: Degree and Centrality"
description: " "
author:
  - name: "Audra Jamai White" 
    url: 
    affiliation: "Umass Amherst - DACSS 679: Social Network Analysis"
    affiliation_url: 
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, Libraries, include=FALSE}
knitr::opts_chunk$set(include = FALSE)
# Document Creation
library(knitr)
library(markdown)
library(yaml)
library(distill)
library(blogdown)
```

## *Briefly describe Smuggler's Cove Drink Index dataset:*

[Smuggler's Cove](https://www.smugglerscovesf.com/#about-section)

:   is a legendary cocktail bar in San Francisco.
    Since opening in 2009, this Tiki Bar has
    developed a dedicated following and received
    many awards. What do people in the restaurant
    industry do when they receive any modicum of
    celebrity?

    *Release a recipe book - of course !*

    It's not like a bar boasting about" featuring
    the largest rum selection in the United
    States, and cocktails from over three
    centuries of rum history" has to worry about
    one of their customers (or competitors) being
    able to replicate their [highly specialized
    bar
    menu](https://www.smugglerscovesf.com/menu).

    The [Smuggler's Cove: Exotic Cocktails, Rum,
    and the Cult of
    Tiki](https://www.smugglerscovesf.com/#book-section)
    book does provide cocktail aficionados and
    home-mixologists with an education in the
    history of rum and "tiki drinks"; along with
    over 80 recipes to inspire our #HomeBarGoals

[Smuggler's Cove Recipe Book Index](https://docs.google.com/spreadsheets/d/1W5eaejaZ9MOu3I36R9TZIus7xlHhkX7JonmBzGHXnTI/edit#gid=942764477)

:   The overlap between [(tech-bros /
    data-nerds) + ((Eldrich) Millennials,
    X-ennials)) + (high class bar flies) ] isn't a
    perfect circle; But it is a very enmeshed Venn
    Diagram!

    Like, any other group creating a 21st social
    nice by relying on pretentiously inaccessible
    hobbies as the basis of social capitol to
    create a hobby to base one's social identity
    on -- there's a sub reddit to share pictures,
    memes, and datasets with like-minded folks.

    [Thanks
    el_joker1](https://www.reddit.com/r/cocktails/comments/5um8zh/smugglers_cove_tiki_cocktail_book_index/)
    !!! I wonder how many other cocktail+data
    geeks have played around with this CVS?

[Network Dataset: Smuggler's Cove Index](https://docs.google.com/spreadsheets/d/1W5eaejaZ9MOu3I36R9TZIus7xlHhkX7JonmBzGHXnTI/edit#gid=942764477)

:   The CSV is a pretty small data set. The
    source.csv contains 6 columns and hundreds of
    rows.

    My network dataset contains 3 columns: Drink
    Name, Drink Ingredient, and Ingredient Type
    (Juice, Rum, Bitters, Fruit, ect).

    These three categories are used to develop and
    Edge list with a small number of redundant
    edge attributes.

# Does a Recipe Book Contain Network Data?

I like to think so.

The network data contained in a cocktail recipe
book can serve as a proxy for some of the more
traditional elements of social network analysis.

![](https://media.giphy.com/media/3o7TKJhutMCl75fkLm/giphy.gif)

Recipes (and recipe meta data) are all a part of
the the larger super-structure of "things humans
eat". Recipe books can be viewed as a codex of
specific elements of specific food cultures[^1].
The elements of recipes in a cookbook have a
relationship to each other because they're are a
part of a curated collection - but that
relationship is defined more by the social facts
of the author's food culture(s) than the
individual decisions.

> *How do the social facts of and author's food
> culture shape their recipe selection?*
>
> The author's of[Smuggler's
> Cove](https://www.smugglerscovesf.com/#about-section)
> could not have opened their bar in a place where
> the community's alcohol food taboo is codified
> in the law, which could be many places - all of
> which are not San Francisco. Since the authors
> did live in California during the early 21st
> century they were exposed to the 21st century
> cocktail culture revival. The ability for the
> authors to write about [Exotic Cocktails, Rum,
> and the Cult of
> Tiki](https://www.amazon.com/Smugglers-Cove-Exotic-Cocktails-Cult/dp/1607747324)
> is wholly contingent upon recreational
> consumption of the alcohol being a deeply
> ingrained element of San Francisco's food
> culture - which developed a cocktail focused
> sub-culture.
>
> When "Prohibition" was a social fact in the
> early 20th century U.S.A. the legally codified
> alcohol taboo was an element of the larger
> culture's class stratification and racial cast
> system. The perpetrators of Temperance Movement
> restricted alcohol access as a means of social
> control - specifically, the socially
> undesirable. While their social betters enjoyed
> the privileged of maintaining (extra)legal
> access to alcohol. This period's cocktail
> culture being determined by access to alcohol
> could be used as a proxy for examining proximity
> to political power during the Temperance
> Movement. Besides the venn diagram of
> communities target by the Temperance Movement
> and the target populations of nascent eugenics
> movement quite possibly being just one circle -
> elements of cocktail culture can serve as a
> proxy to explore network dynamics of early 20th
> pre-globalized international markets, in
> addition to agricultural products immediately
> before The Dust Bowl and advent of
> refrigeration.

#  So What ?

Every time-period has it's own "cocktail culture",
which can be viewed as a network of period-related
products, technological practices, and culinary
aesthetics manifesting in each element of a
cocktail recipe.

I'm curious to find how Drink Names are linked to
each other by sharing ingredients.

##

1.  <div>

    > -   identify initial network format;
    >
    > -   describe and identify the nodes
    >     including:
    >
    >     -   how many nodes are in the dataset;
    >
    > -   what constitutes a tie or edge
    >     including:
    >
    >     -   how many ties,
    >
    >     -   whether ties are
    >         directed/undirected,
    >
    >     -   whether ties are weighted/binary,
    >
    >     -   and how to interpret the value of
    >         the tie if any.
    >
    >     -   Whether or not there are edge
    >         attributes that might be used to:
    >
    >         -   subset data, or
    >         -   stack multiple networks.

    </div>

## Working from the example R syntax:

1.  calculate network density;

2.  degree centrality, and

3.  centralization measures for your network data.

    | Col1 | Col2 | Col3 |
    |------|------|------|
    |      |      |      |
    |      |      |      |
    |      |      |      |

    : Network Maeasure

## *Write up your results briefly:*

-   focus on interpretation of the results.
-   Don't list statistics or provide code,
-   **the goal is to choose a few interesting
    results and provide an interpretation.**

##

## *Provide at least two or three noteworthy results*, including:

1.  the relevant statistics and interpretation.
    For example, explaining which node(s) are most
    central and which are least central.
2.  Discuss (with any related evidence) whether or
    not the node(s) behavior is in line with or
    violates expectations based on the degree
    centrality measure.
3.  What do you make of network density and
    centralization measures?\
4.  What about differences in the binary 100
    million threshold network vs valued trade
    flows network.
5.  Check tutorial for other ideas of results to
    discuss.

> Distill is a publication format for scientific
> and technical writing, native to the web.
>
> Learn more about using Distill for R Markdown at
> <https://rstudio.github.io/distill>.

[^1]: *sFood Culture?*

    The differentiating elements between food
    cultures are what a community eats, how food
    is prepared, when food is prepared, why it's
    prepared a certain way, who prepares the food,
    and who consumes the food. All of which
    combine to create food taboos and culinary
    traditions - the social facts informing a
    foundational aspect of daily life.

    Humans are wild! Throughout time and space we
    have integrated an assortment of plants,
    animals, insects, fungi, and minerals into
    idiosyncratic food cultures. The consumables
    in food culture are particular while the
    technological methods of preparation and
    preservation are shared. Foods are consumed
    raw or cook. Cooked foods can be boiled,
    roasted, baked, or fried. Food preservation
    methods are equally limited. Perishables are
    dried, cured (salt, smoke, sugar, acid, ect.),
    and fermented.[We love to ferment
    things!](https://fermentationassociation.org/fermentation-in-anthropology/)

    Most of my favorite foods are fermented-
    bread, cheese, pickles, booze. Anything can be
    fermented, but fermenting liquids - grain
    mash, plant juice, milk, honey- into beverages
    seems to be universal. The wildest thing and
    our love of fermentation liquids is that the
    end product usually contains a toxin -
    alcohol!

    I'm sure several someone's have researched why
    humans love to create and imbibe intoxicants
    a.k.a poison -[at least we're not the only
    ones.](https://youtu.be/A0Xsqo-n_AE)