### 2022-02-19 docs/3. SNA - Assignments/sna_Week3/Week3_Notebook.Rmd

---
title: "R Notebook"
output: html_notebook
---

```{r setup, Libraries, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)

# Document Creation
library(knitr)
library(markdown)
library(yaml)
library(distill)
library(blogdown)

# Basic Analysis
library(ggplot2)
library(tidyverse)
library(clipr)
library(dbplyr)
library(data.table)
library(stringr)
library(stringi)
library(tibble)
library(tidyr)

# SNA - Social Network Analysis

library(tergm)
library(ergm)
library(network)
library(networkDynamic)
library(networkDynamicData)

library(statnet)
library(igraph)
library(igraphdata)
library(tsna)
library(sna)

```

```{r Set Working Directory, include=FALSE}
getwd()

setwd("C:/Users/white/iCloudDrive/Documents/1_DACSS_UMassAmherst/Dacss_RProjects/3. Sp22_rProjects/Dacss_679- Coursework/GIT-ajw_Learning-SocialNetworkAnalysis/docs/3. SNA - Assignments/sna_Week3")

```

## Basic R: Import & Clean Data

```{r Read raw data into R}

library(readr)

indx_SmugglerCove_recip <- read_csv("C:/Users/white/iCloudDrive/Documents/1_DACSS_UMassAmherst/Dacss_RProjects/3. Sp22_rProjects/Dacss_679- Coursework/GIT-ajw_Learning-SocialNetworkAnalysis/docs/3. SNA-Data/index-by_name-Smuggler_Cove-cocktail_recipies.csv", 
    col_names = FALSE, 
    col_types = cols(X1 = col_skip(),
                     X4 = col_skip()), 
                    na = "empty")

View(indx_SmugglerCove_recip)
summary(indx_SmugglerCove_recip)

 # Smuggler's Cove

smugCovIdx <- indx_SmugglerCove_recip

```

```{r Getting Information on a Dataset:, echo=FALSE}

# list the structure of smugCovIdx
str(smugCovIdx)

# class of an object (numeric, matrix, data frame, etc)
class(smugCovIdx)

# list the variables in smugCovIdx
names(smugCovIdx)

# dimensions of an object
dim(smugCovIdx)
```

```{r & Clean Data}

  library(stringr)
      # Drop NA rows
smugCovIdx <- na.omit(smugCovIdx)

# DATA Cleaning

#     Extract variables
#####

          # extract first word of the column, adds values in new Column at the end of DF

                    #1st pastIt
passIt <- smugCovIdx

colnames(passIt)
                    #2nd passIt
passIt$substring_first <- str_extract(passIt$X5,"(\\w+)")

#####

#####
#     Drop it like...
#####
      # DROP COLUMNS FROM DATA FRAME

                      #3rd PassIt
passIt = subset(passIt, 
                       select = -c(3)
                       )
colnames(passIt)
#####

#####
#     Give Variables a human name
#####

# rename column names with human language

                      # Return it
smugCovIdx <- passIt

names(smugCovIdx )[1] <- "dnkName"
names(smugCovIdx )[2] <- "dnkIng"
names(smugCovIdx )[3] <- "ingType"

colnames(smugCovIdx)
```

#   Project Varriables

    Raw Data:
  indx_SmugglerCove_recip

    DataFrame:
  smugCovIdx
            colname: "dnkName" "dnkIng"  "ingType"
  network_nodes <- smugCovIdx << "dnkName"

    Matrix:
  smug_matrix
  network_edgelist <- smug_matrix

    Network Objects: EDGE LIST
  STATNET: smugEL_sn
  IGRAPH: smugEL_ig

  Edge Vetex:
              dnkName
  Edge Attributes:
                  dnkIng
                  ingType

# Network Info
  Network Objects: EDGE LIST
                  STATNET: smugEL_sn
                  IGRAPH: smugEL_ig
                          ingTypeEL_Dig

```{r EDGELIST: UNDIRECTED Network Objects: statnet & igraph, include=FALSE}

#Lets make some network objects

colnames(smugCovIdx) 
# tell R data should be understood as a matrix
smug_matrix = as.matrix(smugCovIdx,
                        matrix.type="edgelist",
                        directed=FALSE) 
head.matrix(smug_matrix)

network_edgelist <- smug_matrix

  network_nodes <- smugCovIdx 
    network_nodes<- 
      select(network_nodes,
             dnkName)

        ### View(network_nodes)

######
      ######**igraph** 
######
        #igraph object 
network_ig <- 
  graph.edgelist(
    as.matrix(
      smugCovIdx[1:2]), 
    directed=F)
                ## create new edge attribute
  E(network_ig)$dnkIng<-smugCovIdx$dnkIng
  E(network_ig)$ingType<-smugCovIdx$ingType
  #create vertex attributes in igraph
 V(network_ig)[match(V(network_ig)$dnkName,network_nodes$dnkName)]$dnkName<-
   network_nodes$dnkName

      summary(edge_attr(network_ig))

  # head(network_ig)
  # summary(network_ig)
      ######

#####
      #####**statnet** 
#####    
  network_edgelist <- smugCovIdx

  network_nodes <- smugCovIdx 
  network_nodes<- 
    select(network_nodes,
           dnkName)
summary(network_nodes)
View(network_nodes)

  #Now let's create a statnet object
  network_statnet <- network(network_edgelist, 
                             matrix.type = "edgelist",
                             vertex.attr = network_nodes,
                             multiple = T)

  network_statnet%e%'dnkIng'  <-
    network_edgelist$dnkIng

  network_statnet%e%'ingType'<- 
    network_edgelist$ingType

       # list the names: vertex attributes
  view(network_statnet%v% "vertex.names")

summary(network_statnet)
        #####

 ## Network Objects: EDGE LIST
              ##**STATNET:**
  smugEL_sn <- network_statnet

                ##**IGRAPH:** 
  smugEL_ig <- network_ig

```

```{r EDGELIST: IGRAPH: DIRECTED Network: ingType, include=FALSE}

#Lets make some network objects

colnames(smugCovIdx) 
# tell R data should be understood as a matrix
smug_matrix = as.matrix(smugCovIdx,
                        matrix.type="edgelist",
                        directed=FALSE) 
head.matrix(smug_matrix)

network_edgelist <- smug_matrix

  nodes_ingType <- smugCovIdx 
    nodes_ingType<- 
      select(nodes_ingType,ingType)
view(nodes_ingType)
        ### View(nodes_ingType)

######
      ######**igraph** 
######
        #igraph object: ingType
d_network_ig <- 
  graph.edgelist(
    as.matrix(
      smugCovIdx[2:3]), 
    directed=T)

vertex.attributes(d_network_ig)
                ## create new edge attribute
  E(d_network_ig)$dnkIng<-smugCovIdx$dnkIng
  E(d_network_ig)$dnkName<-smugCovIdx$dnkName

  #create vertex attributes in igraph
 V(d_network_ig)[match(V(d_network_ig)$ingType,nodes_ingType$ingType)]$ingType<- nodes_ingType$ingType

   View(d_network_ig)
   head(d_network_ig)
  summary(d_network_ig)
  summary(edge_attr(d_network_ig))
  summary(vertex.attributes(d_network_ig))

   ingTypeEL_Dig <- d_network_ig
```

## Smuggler's Cove Index- NETWORK: EDGELIST: Features & Descriptions

```{r STATNET: Network Description, include=FALSE}

      # class of an object
  class(smugEL_sn )

## access vertex attributes
            # list the names:  vertex attributes
  network::list.vertex.attributes(smugEL_sn)

            # list the names:  edge attributes
  network::list.edge.attributes(smugEL_sn)
      #   summarize: numeric network attribute
  summary(smugEL_sn%v%"dnkName")

## Network Attributes
 print(smugEL_sn)
          # Vertices =  
          # Directed = 
          # hyper = 
          # loops = 
          # multiple = 
          # bipartite = 
          # total edges = 
                # missing edges =
                # non-missing edges = 
          # Vertex attribute names:
          # Edge Attribute Names:

##**NOTE** large quantity of output

  # list the structure of smugCovIdx
# str(smugEL_sn)
```

smugEL_sn

[1] "network"

Network attributes:
  vertices = 274 
  directed = TRUE 
  hyper = FALSE 
  loops = FALSE 
  multiple = TRUE 
  bipartite = FALSE 
  total edges= 672 
    missing edges= 0 
    non-missing edges= 672

 Vertex attribute names: 
    dnkName vertex.names

 Edge attribute names: 
    dnkIng ingType

```{r STATNET: functions that calculate network-level descriptives}
library(statnet)
library(sna)

    #summary of network object, also provides an edge list
summary(smugEL_sn)$no

    #number of nodes in the network
network.size(smugEL_sn)

    #number dyads that exist in network (n*(n-1))
network.dyadcount( smugEL_sn)

    #number of edges present in the network
network.edgecount( smugEL_sn)

    #network density (edge count/dyad count)
gden(smugEL_sn)
  gden(smugEL_sn,
       diag = T,
       mode = "graph")

    ##**diag=* 
    ##boolean indicating whether or not the diagonal should be treated as valid data. 
        ##*Set this true if and only if the data can contain loops. 
        ##*diag is FALSE by default.
    ##**mode="digraph"*
    ##* string indicating the type of graph being evaluated. 
    ##* "digraph" indicates that edges should be interpreted as directed; 
    ##* "graph" indicates that edges are undirected. mode is set to "digraph" by default.

  # grecip calculates the dyadic reciprocity 
grecip(smugEL_sn) 
      #proportion of symmetric dyads
  grecip(smugEL_sn, measure = "dyadic") 
      # returns the ratio of mutuals to non-null dyads.
  grecip(smugEL_sn, measure = "dyadic.nonnull") 
      # the proportion of edges which are reciprocated
  grecip(smugEL_sn, measure = "edgewise") 
      # returns the log of the ratio of the edgewise reciprocity to the density
  grecip(smugEL_sn, measure = "edgewise.lrr") 
      # The correlation measure may be interpreted as the net tendency for edges of similar relative value (with respect to the mean edge value) to occur within the same dyads. 
  grecip(smugEL_sn, measure = "correlation")

      # k-path or k-cycle  compute census statistics (respectively) on one or more input graphs
      ## In addition to aggregate counts of paths or cycles, results may be disaggregated by vertex and co-membership information may be computed.

Usage

  #**kpath* Count paths of length <=3
kpath.census(smugEL_sn, maxlen=3,
             tabulate.by.vertex=FALSE) 
View(kpath.census(smugEL_sn, tabulate.by.vertex=T))

#**kcycle*Count cycles of length <=3
kcycle.census(smugEL_sn, maxlen=3, tabulate.by.vertex=FALSE) 
  View(kcycle.census(smugEL_sn, maxlen=3,tabulate.by.vertex=T))

# counts of cliques by size
clique.census(smugEL_sn, tabulate.by.vertex=F, enumerate=F)
view(clique.census(smugEL_sn, tabulate.by.vertex=TRUE, enumerate=TRUE))

```

gtrans(smugEL_sn)
proportion of transitive triads

symmetrize( smugEL_sn, rule = "weak") 
symmetrize so i<->j iff i->j OR i<-j

symmetrize( smugEL_sn, rule = "strong") 
symmetrize so i<->j iff i->j OR i<-j

dyad.census(smugEL_sn) 
MAN dyad census

triad.census(smugEL_sn) 
standard directed triad census

```{r igraph: Network Description, echo=TRUE}
library(igraph)
      # class of an object
    class(smugEL_ig)

## Network Size

      # Count Vertices
    vcount(smugEL_ig)

      # Count Edges
    ecount(smugEL_ig)

## Network Mode
    # Bipartitie / Unimodal
    is_bipartite(smugEL_ig)

## Edges    
    # Un/directed 
    is_directed(smugEL_ig)

    # weighted or unweighted?
    is_weighted(smugEL_ig)

    # edge weights
    E(smugEL_ig)$weight

## access vertex attributes

    # Generate a list the names of vertex attributes
  V(smugEL_ig)$name

    # access edge attribute  names
  igraph::vertex_attr_names(smugEL_ig)

    # Generate a list the names of edge attributes  
  igraph::edge_attr_names(smugEL_ig)

## Network Attributes
 summary(smugEL_ig)
          # Vertices =  
          # Directed = 
          # hyper = 
          # loops = 
          # multiple = 
          # bipartite = 
          # total edges = 
                # missing edges =
                # non-missing edges = 
          # Vertex attribute names:
          # Edge Attribute Names: 
```
smugEL_ig

[1] "igraph"

Network attributes:
  vertices = 274 
  directed = FALSE
  hyper = FALSE 
  loops = FALSE 
  multiple = TRUE 
  bipartite = FALSE 
  total edges= 672 
    missing edges= 0 
    non-missing edges= 672

 Vertex attribute names: 
    dnkName vertex.names

 Edge attribute names: 
    dnkIng ingType

```{r IGRAPH: DIRECTED Network: ingType, echo=TRUE}

      # class of an object
    class(ingTypeEL_Dig)

## Network Size

      # Count Vertices
    vcount(ingTypeEL_Dig)

      # Count Edges
    ecount(ingTypeEL_Dig)

## Network Mode
    # Bipartitie / Unimodal
    is_bipartite(ingTypeEL_Dig)

## Edges    
    # Un/directed 
    is_directed(ingTypeEL_Dig)

    # weighted or unweighted?
    is_weighted(ingTypeEL_Dig)

    # edge weights
    E(ingTypeEL_Dig)$weight

## access vertex attributes

    # Generate a list the names of vertex attributes
  V(ingTypeEL_Dig)$name

    # access edge attribute  names
  igraph::vertex_attr_names(ingTypeEL_Dig)

    # Generate a list the names of edge attributes  
  igraph::edge_attr_names(ingTypeEL_Dig)

## Network Attributes
 summary(ingTypeEL_Dig)
          # Vertices =  
          # Directed = 
          # hyper = 
          # loops = 
          # multiple = 
          # bipartite = 
          # total edges = 
                # missing edges =
                # non-missing edges = 
          # Vertex attribute names:
          # Edge Attribute Names:

```

  # class of an object
[1] "igraph"

      ## Network Size

  # Count Vertices
[1] 201

  # Count Edges
[1] 672

      ## Network Mode
  # Bipartitie / Unimodal
     is_bipartite
[1] FALSE

    ## Edges    
   # Un/directed 
     is_directed
[1] TRUE

  # weighted or unweighted?
     is_weighted
[1] FALSE

   # edge weights
NULL

#   Week 2: Basic Network Structure: Smuggler's Cove Drink Index
)

## Path Length and Geodesic

### Igraph: Undirected
```{r iGraph: Undirected: Path: Shortest, echo=TRUE}

#     find average shortest path for network
    average.path.length(smugEL_ig,directed=F)

```

```{r iGraph: Undirected: Specific Nodes: Path: Shortest, echo=TRUE}

#     isolating the distances between specific node
    all_shortest_paths(smugEL_ig,"Strozzi","Tornabuoni", weights=NA)$res
## Note: manually tell igraph to ignore edge weights.
```

```{r iGraph: Undirected: Specific Nodes: Path: Length, echo=TRUE}

#     Calculate distances between two nodes
    distances(smugEL_ig,"Bischeri","Ridolfi")

#     Calculate distance between two nodes using unweighted edges
    distances(smugEL_ig,"Bischeri", "Castellani",weights=NA)

```
### Igraph: Directed
```{r iGraph: Directed: Path: Shortest, echo=TRUE}

#     find average shortest path for network
    average.path.length(ingTypeEL_Dig,directed=T)

```

```{r iGraph: Directed:Specific Nodes: Path: Shortest, echo=TRUE}

#     isolating the distances between specific node
    all_shortest_paths(ingTypeEL_Dig,
                       "Strozzi","Tornabuoni", 
                       weights=NA)$res
## Note: manually tell igraph to ignore edge weights.
```

```{r iGraph: Directed: Specific Nodes: Path: Length, echo=TRUE}

#     Calculate distances between two nodes
    distances(ingTypeEL_Dig,"Bischeri","Ridolfi")

#     Calculate distance between two nodes using unweighted edges
    distances(ingTypeEL_Dig,"Bischeri", "Castellani",weights=NA)

```

## Dyad Census

Classifies all dyads in the network as: - 
Reciprocal (mutual), 
or mut - Asymmetric (non-mutual), 
or asym, 
and - Absent, or null

```{r statnet: Dyad Census , echo=TRUE}

    sna::dyad.census(smugEL_sn)

```
plot(sna::dyad.census(smugEL_sn))
```{r iGraph: Undirected: GRAPH: Dyad Census, echo=TRUE}
library(igraph)
graph.density(smugEL_ig)
    ## Density captures, at the macro-level, the proportion of dyads that are present over the possible total number of dyads in the network

      # Generating a random graph for comparison
# ?erdos.renyi.game
      ## In an erdos.renyi.graph, each edge has the same probability of being created. We determine the probability and it returns a random graph with a density that equals (in expectation) this probability.

smugEL_igNod <- vcount(smugEL_ig)

smugEL_igDen <- graph.density(smugEL_ig)

smugEL_igRandoG <- 
  erdos.renyi.game(
    n = smugEL_igNod, 
    p.or.m = smugEL_igDen, 
    directed = F) 
      # where n is the number of nodes, 
      # p.or.m is the probability of drawing an edge, 
      #directed is whether the network is directed or not

#**plot(smugEL_igRandoG, ...** in .R notebook  
     # vertex.size = 5, 
     # vertex.label = 1, 
     # edge.curved = .1, 
     # vertex.color = "tomato", 
     # edge.arrow.size = .1, 
     # edge.width = .5, 
     # edge.color = "grey60")

```

## Triad Census

Classifies all triads in the network. 
The triad census provides a fundamental descriptive insight into the types of triads found in a particular dataset.

```{r statnet: Triad Census, echo=TRUE}

#     Classify all Triad in the network: Undirected Network
   sna::triad.census(smugEL_sn, mode="graph")

#     total number of all four triad types returned by triad census
   sum(sna::triad.census(smugEL_sn, mode="graph"))

```
plot(sna::triad.census(smugEL_sn, mode="graph"))
plot(sum(sna::triad.census(smugEL_sn, mode="graph")))

```{r iGraph: Undirected: Triad Census: Calculating a triad census, echo=TRUE}
#Classify all Triad in the network
   igraph::triad_census(smugEL_ig)

```

 plot(igraph::triad_census(smugEL_ig))

```{r iGraph: Undirected:TRIDADS: Banning triads}

```

## Transitivity or Global Clustering

```{r iGraph: Undirected: weighted/ network transitivity, echo=TRUE}

#     network transitivity:
    transitivity(smugEL_ig)

```

```{r statnet: weighted/ directed network transitivity, echo=TRUE}

#     Directed Networks transitivity:
  gtrans(smugEL_sn)

#   methods for weighted networks, rank and correlation, along with the relevant references.
#  sna::gtrans()

```

## Local Transivity or Clustering
`Local transitivity` (local clustering coefficient), is a technical description of the density of an ego network.
```{r iGraph: Undirected: Transitivity: Network}
#     get global clustering cofficient
  transitivity(smugEL_ig, type="global")

#     get average local clustering coefficient
  transitivity(smugEL_ig, type="average")

```

```{r iGraph: Undirected: Local transitivity: Node List & Ego Network, echo=TRUE}

#     Retrive a list of the vertices we are interested
    V(smugEL_ig)[c("Bischeri" ,
                "Castellani", 
                "Medici",
                "Ridolfi")]

#     check ego network transitivity
transitivity(smugEL_ig,
             type="local", 
             vids=V(smugEL_ig)
             [c("Bischeri",
                "Castellani", 
                "Medici",
                "Ridolfi")]) 
```

## Component Structure and Isolates

```{r Statnet: Isolates, echo=TRUE}

              ## Isolates:

 #List the isolates
 isolates(smugEL_sn)

 # Returns a list of the isolates in the graph or graph set
 isolates(smugEL_sn, diag=FALSE)

 # Is Ego an Isolate?
 is.isolate(smugEL_sn, ego_exp, g=1, diag=FALSE)

 # Adds n isolates to the graph (or graphs)
 add.isolates(smugEL_sn, n_umber, return.as.edgelist = FALSE)

```

```{r iGraph: Undirected: Component Structure, echo=TRUE}
                ## Component Structure:

    # Number of components
 igraph::components(smugEL_ig)$no 
      ##**NOTE:** $no prevents a list of comonents from being printed

# Size of each component
 igraph::components(smugEL_ig)$csize

```

## Network Density

 The density of a network is defined as the proportion of ties present out of all possible ties in a network.

 **Note:**

 the network density command ignores edge values at present.

 If loops are present or there are multiple edges between two nodes, then the results may not make sense.

```{r Statnet: Network Density, echo=TRUE}

            # get network density: with possible loops
   network.density(smugEL_sn)
          ##**WARNING: Network is multiplex - no general way to define density.**

 # get network density **with** loops: statnet
 gden(smugEL_sn, diag=F)

```

 **Igraph** provides an incorrect density calculation as the default assumes  that loops are not included but does not remove them.

 Therefore, the default calculation includes the loops in the numerator but not the denominator of the density calculation.

```{r iGraph: Undirected: network density, echo=TRUE}

    #get network density **with** loops
 network.density(smugEL_ig)
 class(smugEL_ig)
          ##**NOTE** provides an incorrect density calculation 
          ## as the default assumes that loops are not included but does not remove them

```

## Vertex Degree
 **Vertex degree** is a description of an individual node; each vertex in a network may have a different node degree.

 *Node Degree:*
   node degree is one measure of node centrality or popularity; 
 Node degree cis a count of the relationships that a node is involved in or the count of the edges that are incident to a vertex

 **use the network degree command to:**
   *  compute network summary statistics (such as average network degree);
 *  graph the network degree distribution; or
 *  identify particular nodes (such as the five nodes with the most contacts).

```{r Statnet: Vertex Degree, echo=TRUE}

      # statnet: Calculate average network degree

  sna::degree(smugEL_sn)
          # ignores loops or self-nominations

```

```{r iGraph: Undirected: Vertex Degree, echo=TRUE}
# igraph: Calculate average network degree
 # provides node name along with node degree

 # includes loops
  igraph::degree(smugEL_ig)

```

## Network Degree Summary Statistics

```{r Statnet: , echo=TRUE}
      # Basic R: get summary statistics for node attributes
 summary(nodes_ingTypes)
      # provides range, mean, median and standard deviation of a vector.

 # get summary statistics for node degree for the binary network degree

 summary(data.frame(
   name=smugEL_sn%v%"vertex.names",
   totdegree=sna::degree(smugEL_sn),
   indegree=sna::degree(smugEL_sn, cmode="indegree"),
   outdegree=sna::degree(smugEL_sn, cmode="outdegree")
 ))

```

# Centrilization

```{r iGraph: Undirected: DEGREE CENTRALITY: dnkIng}
library(igraph)
library(tidyverse)

# graph the SMUG network

smug_attributes <- smugCovIdx
smug_edj <- smugCovIdx[2:3]
View(smug_edj)
#V1: dnkIng
#V2: ingType

smugNet <- 
  graph.edgelist(as.matrix(smug_edj), directed = T)

V(smugNet)$dnkIng <- 
  smug_attributes$dnkIng[match(V(smugNet)$name, smug_attributes$ingType)]

V(smugNet)$dnkName <- 
  smug_attributes$dnkName[
    match(V(
      smugNet)$name,
      smug_attributes$dnkIng)
    ]
summary(smugNet)

###**PLOT** in .R Doc

```

```{r iGraph: Undirected: DEGREE CENTRALITY: dnkIng}
library(igraph)
library(tidyverse)

# graph the SMUG network

smug_attributes <- smugCovIdx
smug_edj <- smugCovIdx[2:3]
View(smug_edj)
#V1: dnkIng
#V2: ingType

smugNet <- 
  graph.edgelist(as.matrix(smug_edj), directed = T)

V(smugNet)$ingType <- 
  smug_attributes$ingType[match(V(smugNet)$name, smug_attributes$ingType)]

V(smugNet)$dnkName <-
  smug_attributes$dnkName[match(V(smugNet)$name, smug_attributes$dnkName)]

summary(smugNet)

###**PLOT** in .R Doc

```

```{r iGraph: Undirected: DEGREE CENTRALITY: ingType, echo=TRUE}
# graph the SMUG network

smug_attributes <- smugCovIdx
smug_edj <- smugCovIdx[2:3]
View(smug_edj)
#V1: dnkIng
#V2: ingType

smugNet <- 
  graph.edgelist(as.matrix(smug_edj), directed = T)

V(smugNet)$dnkName <-
  smug_attributes$dnkName[match(V(smugNet)$name, smug_attributes$dnkIng)]

summary(smugNet)
View(smugNet)

###**PLOT** in .R Doc

```

```{r Statnet: Visualization, echo=TRUE}

 # add in vertex labels and change the size and color of these labels
 gplot(smugEL_sn, 
       displaylabels=TRUE,
       label.cex=.75, 
       label.col="black"
 )

 # For directed graphs, you can turn off the arrows
 gplot(smugCovIdx, 
       displaylabels=TRUE, 
       label.cex=.75, 
       label.col="black", 
       usearrows = FALSE)

 # differentiate nodes based on their attributes
 gplot(smugEL_sn, 
       displaylabels=TRUE,
       label.cex=.75,
       label.col="black",
       vertex.col=att$dnkName
 )

 legend("bottomleft",
        fill=0:4,
        legend=paste("DEPT",0:4),
        cex=0.75
 )

 # changed the shape of the nodes based on the level of their positions. 
 ## You change the shape of the nodes by specifying the number of sides you want the shape to have using the vertex.sides argument. 
 gplot(smugEL_sn, 
       displaylabels=TRUE,
       label.cex=.75,
       label.col="black",
       vertex.cex = 2,
       vertex.sides=(att$LEVEL+2)
 )

```

```{r Statnet: , echo=TRUE}

```

## Visualization: ggplot2
```{r Visualization: ggplot2}

# Dependencies
 library(network)
 library(sna)
 library(ggplot2)

library(visNetwork)
library(threejs)
 library(networkD3)
 library(ndtv)
 # ggnetwork - This package allows to pass network objects to ggplot2 and provides geometries to plot their elements.

 library(ggnetwork)

  # visualize your network graphs 
 library(GGally)

## [network visualization with ggplot2]

```

#
```{r Statnet: , echo=TRUE}

```

```{r igraph: , echo=TRUE}

```

Distill is a publication format for scientific
and technical writing, native to the web.

Learn more about using Distill for R Markdown at
<https://rstudio.github.io/distill. ##

# iGraph Directed Network Object

## Degree Centrality
We can break large social networks down into their constituent parts. These constituent parts are referred to as “motifs”. The most basic motif consists of two nodes and is called a dyad. Edges in a network signify the presence or absence of dyadic relations. It follows that a dyad in an undirected network can have two unique configurations: connected or disconnected; and three unique configurations in a directed network (mutual, assymetric, and null

```{r iGraph: Directed: GRAPH: Dyad Census, echo=TRUE}

graph.density(ingTypeEL_Dig)
    ## Density captures, at the macro-level, the proportion of dyads that are present over the possible total number of dyads in the network

      # Generating a random graph for comparison
# ?erdos.renyi.game
      ## In an erdos.renyi.graph, each edge has the same probability of being created. We determine the probability and it returns a random graph with a density that equals (in expectation) this probability.

ingTypeEL_DigNod <- vcount(ingTypeEL_Dig)

ingTypeEL_DigDen <- graph.density(ingTypeEL_Dig)

smugEL_DigRandoG <- 
  erdos.renyi.game(
    n = smugEL_igNod, 
    p.or.m = smugEL_igDen, 
    directed = T) 
      # where n is the number of nodes, 
      # p.or.m is the probability of drawing an edge, 
      #directed is whether the network is directed or not

#**plot(smugEL_DigRandoG, ...** in .R notebook  
     # vertex.size = 5, 
     # vertex.label = 1, 
     # edge.curved = .1, 
     # vertex.color = "tomato", 
     # edge.arrow.size = .1, 
     # edge.width = .5, 
     # edge.color = "grey60")

```

```{r iGraph: Directed: Degree Centrality}
# Degree centrality is calculated using the degree function in R. It returns how many edges each node has.
view(degree(ingTypeEL_Dig))

ingT_deg <- ingTypeEL_Dig
V(ingT_deg)$degree <- degree(ingT_deg) # assignment

# plot(ingT_deg, 
#     vertex.label.cex = .6, 
#     vertex.label.color = "black", 
#     vertex.size = V(ingTypeEL_Dig)$degree, 
#     vertex.label.cex = .2) # sized by degree

```

```{r iGraph: Directed: Betweenness Centrality}

view(betweenness(ingTypeEL_Dig, directed = FALSE))

# view(betweenness(ingTypeEL_Dig, directed = TRUE))
              # Yeilds 0

ingTypeEL_Between <- ingTypeEL_Dig
# We can again assign the output of betweenness() to a variable in the network and size the nodes according to it.

V(ingTypeEL_Between)$betweenness <- 
  betweenness(ingTypeEL_Between, 
              directed = F)

# plot(ingTypeEL_Between, 
#     vertex.label.cex = .6, 
#     vertex.label.color = "black", 
#     vertex.size = V(ingTypeEL_Between)$betweenness) # sized by betweenness

```

With closeness centrality we again make use of the shortest paths between nodes. We measure the distance between two nodes as the length of the shortest path between them. Farness, for a given node, is the average distance from that node to all other nodes. Closeness is then the reciprocal of farness (1/farness).
```{r iGraph: UnDirected:  Closeness Centrality}

view(closeness(smugEL_ig))

  # We assign it to a node variable and plot the network, adjusting node size by closeness.
ingTyp_Cent <- smugEL_ig #**UNDIRECTED **

V(ingTyp_Cent)$closeness <- closeness(ingTyp_Cent)

View(ingTyp_Cent)

# plot(ingTyp_Cent,
#     vertex.label.cex = .6, 
#     vertex.label.color = "black", 
#     vertex.size = V(ingTyp_Cent)$closeness/max(V(ingTyp_Cent)$closeness) * 20)

```

Degree centrality only takes into account the number of edges for each node, but it leaves out information about ego’s alters
```{r iGraph: Directed: Eigenvector Centrality}

view(evcent(ingTypeEL_Dig)$vector)
# view(evcent(smugEL_ig)$vector)
ingTyp_Vector <- ingTypeEL_Dig
V(ingTyp_Vector)$eigenvector <- evcent(ingTyp_Vector)$vector

# plot(ingTyp_Vector,
#     vertex.label.cex = .6, 
#     vertex.label.color = "black", 
#     vertex.size = V(ingTyp_Vector)$eigenvector/max(V(ingTyp_Vector)$eigenvector) * 20)
```

 “Interpretively, the Bonacich power measure corresponds to the notion that the power of a vertex is recursively defined by the sum of the power of its alters. The nature of the recursion involved is then controlled by the power exponent: positive values imply that vertices become more powerful as their alters become more powerful (as occurs in cooperative relations), while negative values imply that vertices become more powerful only as their alters become weaker (as occurs in competitive or antagonistic relations).”
```{r iGraph: Directed: Bonacich Centrality}

ingTyp_Cent <- ingTypeEL_Dig

V(ingTyp_Cent)$bonacich <- 
  power_centrality(ingTyp_Cent, exponent = -2, rescale = T)

V(ingTyp_Cent)$bonacich <- 
  ifelse(V(ingTyp_Cent)$bonacich < 0, 0, V(ingTyp_Cent)$bonacich)
```

```{r iGraph: Directed: Page Rank}

ingTyp_Cent <- ingTypeEL_Dig

V(ingTyp_Cent)$page_rank <- page_rank(ingTyp_Cent, directed = TRUE)$vector

```

 plot(ingTyp_Cent,
     vertex.label.cex = .6, 
     vertex.label.color = "black", 
     vertex.size = V(ingTyp_Cent)$page_rank/max(V(ingTyp_Cent)$page_rank) * 20)

```{r FUCK IT: Measure Correlations, eval=FALSE, include=FALSE}
# extract all the vertex attributes
all_atts <- 
  lapply(list.vertex.attributes(ingTyp_Cent),function(x) get.vertex.attribute(ingTyp_Cent,x))

# bind them into a matrix
all_atts <- 
  do.call("cbind", all_atts)

# add column nams
colnames(all_atts) <- 
  list.vertex.attributes(ingTyp_Cent)

# drop the family variable
all_atts <- 
  data.frame(all_atts[,2:ncol(all_atts)])

# convert all to numeric
all_atts <- 
  sapply(all_atts, as.numeric)

# produce a correlation matrix
cormat <- 
  cor(all_atts)

# melt it using reshape to function melt() to prepare it for ggplot which requires long form data
setDT(as.data.frame(cormat))
melted_cormat <- melt(cormat)

ggplot(data = melted_cormat, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile() +
  scale_fill_distiller(palette = "Spectral", direction=-2) +
  xlab("") +
  ylab("")
```

```{r FUCKIT : Centralization and Degree Distributions, eval=FALSE, include=FALSE}
# extract all the vertex attributes
all_atts <- lapply(list.vertex.attributes(ingTyp_Cent),function(x) get.vertex.attribute(ingTyp_Cent,x))
# bind them into a matrix
all_atts <- do.call("cbind", all_atts)
# add column nams
colnames(all_atts) <- list.vertex.attributes(ingTyp_Cent)
# drop the family variable
all_atts <- data.frame(all_atts[,2:ncol(all_atts)])
# convert all to numeric
all_atts <- sapply(all_atts, as.numeric)
# produce a correlation matrix
cormat <- cor(all_atts)

as.data.table(cormat)

view()
# melt it using reshape to function melt() to prepare it for ggplot which requires long form data
melted_cormat <- melt(cormat)
ggplot(data = melted_cormat, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile() +
  scale_fill_distiller(palette = "Spectral", direction=-2) +
  xlab("") +
  ylab("")
```