### 2022-02-22 docs/0. Setting UP Project/Week0_About_Distill-Blog.Rmd

---
title: "Social Network Analysis: Week 0: About this dataset"
description: " "
author:
  - name: "Audra Jamai White" 
    url: 
    affiliation: "UMass Amherst: DACSS 679: Social Network Analysis: Assignments"
    affiliation_url: 
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, Libraries, include=FALSE}
knitr::opts_chunk$set(include = FALSE)
# Document Creation
library(knitr)
library(markdown)
library(yaml)
library(distill)
library(blogdown)
```

## *About the Smuggler's Cove Drink Index dataset:*

[Smuggler's Cove](https://www.smugglerscovesf.com/#about-section)

:   is a legendary cocktail bar in San Francisco.
    Since opening in 2009, this Tiki Bar has
    developed a dedicated following, received many
    awards, and is high-key famous among cocktail
    corps d'elite.

![](https://media.giphy.com/media/l2JJIQmEvS4hBbnjO/giphy.gif)

[Exotic Cocktails, Rum, and the Cult of Tiki](https://www.smugglerscovesf.com/#book-section)

:   What do people in the restaurant industry do
    when they receive any modicum of celebrity?

    *Release a recipe book - of course !*

    It's not like a bar boasting about" featuring
    the largest rum selection in the United
    States, and cocktails from over three
    centuries of rum history" has to worry about
    one of their customers (or competitors) being
    able to replicate their [highly specialized
    bar
    menu](https://www.smugglerscovesf.com/menu).

    The [Smuggler's Cove: Exotic Cocktails, Rum,
    and the Cult of
    Tiki](https://www.smugglerscovesf.com/#book-section)
    book does provide cocktail aficionados and
    home-mixologists with an education in the
    history of rum and "tiki drinks"; along with
    over 80 recipes to inspire our #HomeBarGoals

[Smuggler's Cove Recipe Book Index](https://docs.google.com/spreadsheets/d/1W5eaejaZ9MOu3I36R9TZIus7xlHhkX7JonmBzGHXnTI/edit#gid=942764477)

:   The overlap between [(tech-bros /
    data-nerds) + ((Eldrich) Millennials,
    X-ennials)) + (high class bar flies) ] isn't a
    perfect circle; But it is a very enmeshed Venn
    Diagram!

    Like, any other group creating a 21st social
    nice by relying on pretentiously inaccessible
    hobbies as the basis of social capitol to
    create a hobby to base one's social identity
    on -- there's a sub reddit to share pictures,
    memes, and datasets with like-minded folks.

    [Thanks
    el_joker1](https://www.reddit.com/r/cocktails/comments/5um8zh/smugglers_cove_tiki_cocktail_book_index/)
    !!! I wonder how many other cocktail+data
    geeks have played around with this CVS?

[Network Dataset: Smuggler's Cove Index](https://docs.google.com/spreadsheets/d/1W5eaejaZ9MOu3I36R9TZIus7xlHhkX7JonmBzGHXnTI/edit#gid=942764477)

:   The CSV is a pretty small data set. The
    source.csv contains 6 columns and hundreds of
    rows.

    My network dataset contains 3 columns: Drink
    Name, Drink Ingredient, and Ingredient Type
    (Juice, Rum, Bitters, Fruit, ect).

    These three categories are used to develop and
    Edge list with a small number of redundant
    edge attributes.

# Does a Recipe Book Contain Network Data?

I like to think so.

Every time-period has it's own "cocktail culture",
which can be viewed as a network of period-related
products, technological practices, and culinary
aesthetics manifesting in each element of a
cocktail recipe.

Converting the data and metadata in recipe book
into network data could serve as a proxy to
reflect on more conventional network datasets.
Researchers into business and profitability could
recipe books cum network data to perform
historical market research - but all of those
people are not me.

![](https://media.giphy.com/media/3o7TKJhutMCl75fkLm/giphy.gif)

Recipes (and recipe meta data) are all a part of
the the larger super-structure of "things humans
eat". Recipe books can be viewed as a codex of
specific elements of specific food cultures[^1].
The elements of recipes in a cookbook have a
relationship to each other because they're are a
part of a curated collection - but that
relationship is defined more by the social facts
of the author's food culture(s) than the
individual decisions.

How do the social facts of and author's food culture shape their recipe selection?

:   The author's of[Smuggler's
    Cove](https://www.smugglerscovesf.com/#about-section)
    could not have opened their bar in a place
    where the community's alcohol food taboo is
    codified in the law, which could be many
    places - all of which are not San Francisco.
    Since the authors did live in California
    during the early 21st century they were
    exposed to the 21st century cocktail culture
    revival. The ability for the authors to write
    about [Exotic Cocktails, Rum, and the Cult of
    Tiki](https://www.amazon.com/Smugglers-Cove-Exotic-Cocktails-Cult/dp/1607747324)
    is wholly contingent upon recreational
    consumption of the alcohol being a deeply
    ingrained element of San Francisco's food
    culture - which developed a cocktail focused
    sub-culture.

    When "Prohibition" was a social fact in the
    early 20th century U.S.A. the legally codified
    alcohol taboo was an element of the larger
    culture's class stratification and racial cast
    system. The perpetrators of Temperance
    Movement restricted alcohol access as a means
    of social control - specifically, the socially
    undesirable. While their social betters
    enjoyed the privileged of maintaining
    (extra)legal access to alcohol. This period's
    cocktail culture being determined by access to
    alcohol could be used as a proxy for examining
    proximity to political power during the
    Temperance Movement. Besides the venn diagram
    of communities target by the Temperance
    Movement and the target populations of nascent
    eugenics movement quite possibly being just
    one circle - elements of cocktail culture can
    serve as a proxy to explore network dynamics
    of early 20th pre-globalized international
    markets, in addition to agricultural products
    immediately before The Dust Bowl and advent of
    refrigeration.

# So What ?

Cocktails are build with specific ingredients,
methods, and presentations to fit contemporary
cocktail culture aesthetics.

![](https://media3.giphy.com/media/zbW5zoQetefZwdBR6Z/200w.webp?cid=ecf05e477of2cy4i1vdat12zy8n4ufrcuuikyv6qxkmwe1z2&rid=200w.webp&ct=g){width="400"}

There is not an organization with the self-imposed
responsibility and authority to standardize the
recopies and classification of cocktails while
maintain records attributing a drinks development
to a specific bar or bartender. Can we use SNA
methods to suss out anecdotal aspects of
contemporary cocktail culture?

To start: I'm curious to see what sort of
relationships exists between Tiki Cocktail
Ingredients? Cocktails are build with specific
ingredients, methods, and presentations to fit
contemporary cocktail culture aesthetics.

*For example:* "gin goes with lime (not lemon);"
but why is that? *Short answer:* European
imperialism and systematic wealth extraction of
the global south. It's the same reason why my
Jamaican Grandmother said "kidney beans go in rice
& peas (not English peas)"

*Another example:* The fictional James Bond being
cited as an authoritative voice on whether a
proper Martini's should be shaken or stirred.
*Short answer:* Cocktails with eggs and milk are
shaken to create foam. Spirit forward drinks tend
to be stirred to limit dilution. *You can shake a
drink with carbonated ingredients if you want to.*

# What makes a Tiki Drink a "Tiki Drink?"

This isn't the "ask 10 people and get 10 different
answers" sort of question, but all of the answers
are based on each person's professionally
(un)informed opinion. As a former bartender - I
believe in the authority of the published opinions
of professional bartenders and [cocktail
historians](https://www.theatlantic.com/health/archive/2010/06/how-to-be-a-drink-historian/58724/).

Their opinion is based in direct experience and
observation of the subject in question. Which is
the sort of thing empiricists do, but since they
are building cocktails instead of a dataset, their
answers are given the same consideration of
qualitative research.

## To start:

I'm curious to see what sort of relationships
exists between cocktail ingredients?

![](C:/Users/white/AppData/Local/RStudio/tmp/paste-FF879F11.png){width="500"}

> `D`istill is a publication format for scientific
> and technical writing, native to the web.
>
> Learn more about using Distill for R Markdown at
> <https://rstudio.github.io/distill>.

[^1]: *sFood Culture?*

    The differentiating elements between food
    cultures are what a community eats, how food
    is prepared, when food is prepared, why it's
    prepared a certain way, who prepares the food,
    and who consumes the food. All of which
    combine to create food taboos and culinary
    traditions - the social facts informing a
    foundational aspect of daily life.

    Humans are wild! Throughout time and space we
    have integrated an assortment of plants,
    animals, insects, fungi, and minerals into
    idiosyncratic food cultures. The consumables
    in food culture are particular while the
    technological methods of preparation and
    preservation are shared. Foods are consumed
    raw or cook. Cooked foods can be boiled,
    roasted, baked, or fried. Food preservation
    methods are equally limited. Perishables are
    dried, cured (salt, smoke, sugar, acid, ect.),
    and fermented.[We love to ferment
    things!](https://fermentationassociation.org/fermentation-in-anthropology/)

    Most of my favorite foods are fermented-
    bread, cheese, pickles, booze. Anything can be
    fermented, but fermenting liquids - grain
    mash, plant juice, milk, honey- into beverages
    seems to be universal. The wildest thing and
    our love of fermentation liquids is that the
    end product usually contains a toxin -
    alcohol!

    I'm sure several someone's have researched why
    humans love to create and imbibe intoxicants
    a.k.a poison -[at least we're not the only
    ones.](https://youtu.be/A0Xsqo-n_AE)