### 2022-02-22 docs/3. SNA - Assignments/Week3_DRAFT_Results.Rmd

---
title: "Social Network Analysis: Week 3: Degree and Centrality"
description: " "
author:
  - name: "Audra Jamai White" 
    url: 
    affiliation: "Umass Amherst - DACSS 679: Social Network Analysis"
    affiliation_url: 
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, Libraries, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Document Creation
library(knitr)
library(markdown)
library(yaml)
library(distill)
library(blogdown)

# Basic Analysis
library(ggplot2)
library(tidyverse)
library(clipr)
library(dbplyr)
library(data.table)
library(stringr)
library(stringi)
library(tibble)
library(tidyr)

# SNA - Social Network Analysis

library(tergm)
library(ergm)
library(network)
library(networkDynamic)
library(networkDynamicData)

library(statnet)
library(igraph)
library(igraphdata)
library(tsna)
library(sna)

```

## Working from the example R syntax:

1.  calculate network density;
2.  degree centrality, and
3.  centralization measures for your network data.

## *Write up your results briefly:*

-   focus on interpretation of the results.
-   Don't list statistics or provide code,
-   **the goal is to choose a few interesting
    results and provide an interpretation.**

## *Briefly describe the dataset:*

# 1. identify initial network format

<div>

> -   identify initial network format;
>
>     -   **Smuggler's Cove Index was converted
>         into an edgelist**
>
> -   what constitutes a tie or edge including:
>
>     -   Ties
>
>         -   Number:
>
>         -   Directed:
>
>         -   Binary:
>
>     -   interpret the value of the tie if any.
>
>     -   edge attributes that might be used to:
>
>     <!-- -->
>
>         -   subset data, or
>         -   stack multiple networks.

</div>

# 2. describe and identify the nodes including:

<div>

**Smuggler's Cove Edge-list contains:**

> -   Nodes: 274
>
>     -   Nodes Values:
>
>         -   [Drink Name]{.smallcaps}
>
>         -   [Drink Ingredient]{.smallcaps}
>
> -   Edges: 672
>
>     -   Drink Ingredient Type

</div>

# 3. what constitutes a tie or edge including:

-   how many ties,

-   whether ties are directed/undirected,

-   whether ties are weighted/binary,

-   and how to interpret the value of the tie if
    any.

-   Whether or not there are edge attributes that
    might be used to:

    -   subset data, or
    -   stack multiple networks.

# *Provide at least two or three noteworthy results*, including:

## 1. the relevant statistics and interpretation. For example, explaining which node(s) are most central and which are least central.

## 2. Discuss (with any related evidence) whether or not the node(s) behavior is in line with or violates expectations based on the degree centrality measure.

## 3.What do you make of network density and centralization measures?

## 4. What about differences in the binary 100 million threshold network vs valued trade flows network.

## 5. Check tutorial for other ideas of results to discuss.

> Distill is a publication format for scientific
> and technical writing, native to the web.
>
> Learn more about using Distill for R Markdown at
> <https://rstudio.github.io/distill>.