### 2022-02-25 docs/3. SNA - Assignments/sna_Week3/week3_Plots/Scripts_week3_Plots.Rmd

# IGRAPH
```{r igraph:}

library(igraph)

```

# Plotting: network data

###**Plot Name: smugIG_plot-R_blue-curved-small*
```{r igraph:plot: basic R }

library(igraph)
plot(smugEL_ig, 
     vertex.size=3,
     vertex.color = "blue",
     vertex.frame.color = NA,
     vertex.label = NA,
     edge.curved = 1.7,
     edge.arrow.size = .3, 
     edge.width = .5)

```

###**Plot: Kamada Kawai **
    ###**Plot Name** igraph_kamadaLayout_smug
```{r igraph: Plot: Kamada Kawai}

library(igraph)

 # run the layout function on our graph

 kamadaLayout <-
   layout.kamada.kawai(smugEL_ig)

# change the default layout setting to equal the layout we generated above
 plot(smugEL_ig, 
      layout = kamadaLayout, 
      vertex.size = 3, 
      vertex.color = "black", 
      vertex.frame.color = NA, 
      vertex.label.cex = .7,  
      vertex.label = NA, 
      edge.curved = .3, 
      edge.arrow.size = .1, 
      edge.width = .5)
```

###**Plot: Frucherman-Reingold **
   ###**Plot Name** igrap-smugEL_Frucherman-Reingold
```{r igraph:}

library(igraph)

# run the layout function on our graph
 fruchtermanLayout <- 
   layout.fruchterman.reingold(smugEL_ig)

 # and then we change the default layout setting to equal the layout we generated above

 plot(smugNet, 
      vertex.size = 6, 
      vertex.color= "orange",
      vertex.label.cex = .5, 
      vertex.label.color = "black", 
      edge.with = .5,
      edge.curved = 5,
      edge.arrow.size = 0.1)

```

# Plotting: attribute data
###**Plot: Color & Attribute**
              ###**Plot Name: $Vertex Naame and $color **

```{r igraph:}

library(igraph)
 V(smugEL_ig)$color <- ifelse(V(smugEL_ig)$name == "PUKA PUNCH", "red", "yellow")
  V(smugEL_ig)$color <- NA
  V(smugEL_ig)$color <- ifelse(V(smugEL_ig)$name == "RUMBUSTION PUNCH", "red","yellow")
  V(smugEL_ig)$color <- ifelse(V(smugEL_ig)$name == "TRIUMVIRATE PUNCH", "red", V(smugEL_ig)$color)
  V(smugEL_ig)$color <- ifelse(V(smugEL_ig)$name == "KAHIKO PUNCH", "red", V(smugEL_ig)$color)
  V(smugEL_ig)$color <- ifelse(V(smugEL_ig)$name == , "black", V(smugEL_ig)$color)
  V(smugEL_ig)$color <- ifelse(V(smugEL_ig)$name ==  "DOCTOR FUNK" , "green",V(smugEL_ig)$color)
  V(smugEL_ig)$color <- ifelse(V(smugEL_ig)$name ==  "AKU AKU", "green",V(smugEL_ig)$color)
  V(smugEL_ig)$color <- ifelse(V(smugEL_ig)$name ==  "JET PILOT" , "green",V(smugEL_ig)$color)
  V(smugEL_ig)$color <- ifelse(V(smugEL_ig)$name == "SINGAPORE SLING", "purple", V(smugEL_ig)$color)
  V(smugEL_ig)$color <- ifelse(V(smugEL_ig)$name ==  "SCORPION", "purple", V(smugEL_ig)$color)
  V(smugEL_ig)$color <- ifelse(V(smugEL_ig)$name == "MAI TAI", "purple", V(smugEL_ig)$color)
  V(smugEL_ig)$color <- ifelse(V(smugEL_ig)$name == "MEXICAN EL DIABLO", "black", V(smugEL_ig)$color)

      plot(smugEL_ig, 
           vertex.size = 3, 
           vertex.frame.color =  "black", 
           vertex.label.cex = .7, 
           vertex.label = NA, 
           edge.curved = .5, 
           edge.arrow.size = .3)

```

```{r igraph:}

library(igraph)

```

# Statnet
```{r STATNET:}

library(statnet)
library(sna)

```

# Plotting: network data
# Plotting: attribute data