### 2022-03-06 docs/3. SNA - Assignments/DRAFT_Week3-Distill-Blog_Degree-and-Centrality.Rmd
3.6.22

---
title: "Social Network Analysis: Week 3: Degree and Centrality"
description: " "
author:
  - name: "Audra Jamai White" 
    url: 
    affiliation: "Umass Amherst - DACSS 679: Social Network Analysis"
    affiliation_url: 
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, Libraries, include=FALSE}
knitr::opts_chunk$set(include = FALSE)
# Document Creation
library(knitr)
library(markdown)
library(yaml)
library(distill)
library(blogdown)
```

![](https://media.giphy.com/media/3ornkcUTDQyU4E6kRa/giphy.gif)

![](https://media.giphy.com/media/3ornjK3aalyWAEhxDi/giphy.gif)

![](https://media.giphy.com/media/3o6Yg4Hgd2bt4802xW/giphy.gif)

![](https://media.giphy.com/media/l2JJzfwyTZYUbjQn6/giphy.gif)

# [Research Question:]{.smallcaps}

## Can we use SNA methods to suss out anecdotal aspects of contemporary cocktail culture?

Probably!

The difficulty lies in finding the right type of
data and formatting it in such a way to answer
meaningful questions.

![](https://media.giphy.com/media/l44QthVdK0aN0h0BO/giphy.gif)

[#ButAlso]{.smallcaps} Determining what those
*meaningful questions* are*.*

### Does a Recipe Book Contain Network Data?

The elements of recipes in a cookbook have a
relationship to each other because they're are a
part of a curated collection - but that
relationship is defined more by the social facts
of the author's food culture(s) than the
individual decisions.

![](https://c.tenor.com/lNr4hoGtM3cAAAAC/thomas-jane-josephus-miller.gif)

Recipes (and recipe meta data) are all a part of
the the larger super-structure of "things humans
eat". Each recipe book can serve as a topical
codex for author's food cultures[^1].

[Smuggler's Cove](https://www.smugglerscovesf.com/#about-section)

:   is a legendary cocktail bar in San Francisco.
    Since opening in 2009, this Tiki Bar has
    developed a dedicated following and received
    many awards.

Source Text: [Smuggler's Cove: Exotic Cocktails, Rum, and the Cult of Tiki](https://www.smugglerscovesf.com/#book-section)

:   Fatures 80 cocktail recipes accompanied by the
    history of rum and tiki drinks.

Source Data: [Smuggler's Cove: Recipe Book Index](https://docs.google.com/spreadsheets/d/1W5eaejaZ9MOu3I36R9TZIus7xlHhkX7JonmBzGHXnTI/edit#gid=942764477)

:   [Thanks
    el_joker1](https://www.reddit.com/r/cocktails/comments/5um8zh/smugglers_cove_tiki_cocktail_book_index/)
    for converting the index into a CSV and
    posting on Reddit

    Smuggler's Cove CSV is a pretty small data
    set. The source.csv contains 6 columns and
    several hundred rows.

Network Data: Smuggler's Cove Index

:   My network dataset contains 3 columns:

    -   [Drink Name]{.smallcaps}

    -   [Drink Ingredient]{.smallcaps}
        ^Lime Juice/Rye/Orgeat/Lime Wedge^

    -   [Ingredient Type]{.smallcaps}
        ^Juice/ Whiskey/ Syrup/Fruit^

    These three categories are used to develop and
    Edge list with a small number of redundant
    edge attributes.

# What makes a Tiki Drink a "Tiki Drink?"

This isn't the "ask 10 people and get 10 different
answers" sort of question - but all of the answers
are based on each person's (un)informed opinion
and not a professional organization [issuing
product
standards](https://www.gayot.com/wine/wine-map/france.html)
or establishing professional standards with
data-driven research.

![](https://media.giphy.com/media/3oz8xxBsDMZWcMCHoQ/giphy.gif)

## To start:

### I'm curious to see what sort of relationships exist between the components of [Smuggler's Cove](https://www.smugglerscovesf.com/#about-section) Tiki Cocktail Recipies?

# ## \*Write up your results briefly: \*

\* focus on interpretation of the results.

\* Don't list statistics or provide code,

\* \*\*the goal is to choose a few interesting
results and provide an interpretation.\*\*

## \*Briefly describe the dataset:\*

Smuggler's Bay Drink Recipe Book Index

:   Is an edge-list with one edge attribute:

    Vertices:

    -   Drink Name

    -   Drink Ingredient

    Edge Attribute:

    -   Drink Ingredient Type

Network Description

:   Nodes:

    Edges:

    Ties: Are a connection between Drink Names and
    Drink Ingredients. The network is undirected
    and unweighted. Undirected: because the
    relationship between Drink Names and Drink
    Ingredients lacks any sort of power dynamic.
    Edge weights could be added by including the
    amount of each ingredient required for each
    drink recipe.

    -   Statnet: number, directed, unweighted

    -   igraph: number, udirected, unweighted

    Smuggler's Bay Edglist is long but
    uncomplicated: which means there are few ways
    to subset data to help answer our research
    question: ***what sort of relationships exist
    between the components ?*** Which makes Drink
    Ingredient Type edge attribute very useful.

- Whether or not there are edge attributes

that might be used to:

- subset data, or

- stack multiple networks.

## \*Provide at least two or three noteworthy

results\*, including:

1\. the relevant statistics and interpretation.

For example, explaining which node(s) are most
central and which are least central.

2\. Discuss (with any related evidence) whether or
not the node(s) behavior is in line with or
violates expectations based on the degree
centrality measure.

3\. What do you make of network density and
centralization measures?

4\. What about differences in the binary 100
million threshold network vs valued trade flows
network.

5.  Check tutorial for other ideas of results to
    discuss.

    *Provide at least two or three noteworthy
    results*, including:

<!-- -->

1.  the relevant statistics and interpretation.

    -   For example, explaining which **node(s)
        are most central and which are least
        central.**

2.  Discuss (with any related evidence) whether or
    not the **node(s) behavior is in line with or
    violates expectations based on the degree
    centrality measure.**

    1.  **What do you make of network density and
        centralization measures?**

> Distill is a publicacien is a p ublicanttecnical
> witing, native to the web.
>
> Learn more about using Distill for R Markdown at
> <https://rstudio.github.io/distill>.

[^1]: *sFood Culture?*

    The differentiating elements between food
    cultures are what a community eats, how food
    is prepared, when food is prepared, why it's
    prepared a certain way, who prepares the food,
    and who consumes the food. All of which
    combine to create food taboos and culinary
    traditions - the social facts informing a
    foundational aspect of daily life.

    Humans are wild! Throughout time and space we
    have integrated an assortment of plants,
    animals, insects, fungi, and minerals into
    idiosyncratic food cultures. The consumables
    in food culture are particular while the
    technological methods of preparation and
    preservation are shared. Foods are consumed
    raw or cook. Cooked foods can be boiled,
    roasted, baked, or fried. Food preservation
    methods are equally limited. Perishables are
    dried, cured (salt, smoke, sugar, acid, ect.),
    and fermented.[We love to ferment
    things!](https://fermentationassociation.org/fermentation-in-anthropology/)

    Most of my favorite foods are fermented-
    bread, cheese, pickles, booze. Anything can be
    fermented, but fermenting liquids - grain
    mash, plant juice, milk, honey- into beverages
    seems to be universal. The wildest thing and
    our love of fermentation liquids is that the
    end product usually contains a toxin -
    alcohol!

    I'm sure several someone's have researched why
    humans love to create and imbibe intoxicants
    a.k.a poison -[at least we're not the only
    ones.](https://youtu.be/A0Xsqo-n_AE)